Succinct representations for abstract interpretation

### 2011-11-01 papers/PLDI2012/implicitization.tex
Remarques et typos sur les dernières pages

@@

@@

SSA introduces $\phi$-functions at the head of a control code to define variables whose value depends on which incoming edge to this control node was last taken. For instance, for \lstinline|if (...) { x = 2*x+1; } else { x= 0; }|, then $x_2$ is defined as $\phi(2x_1+1,0)$. In this framework, each variable is uniquely defined as an arithmetic ($+$, $-$, $\times$, $/$) function of other variables that themselves cannot be represented as affine linear functions, because they are defined using $\phi$-functions, bitwise arithmetic, loads from memory, or return values from function calls.

@@

\begin{figure}[h]
  \begin{center}
    \input{./techniques.tex}
  \end{center} 
  \vspace{-20px}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).
  The $\subsetneq$ bars gives the percentage of invariants stronger (more precise; smaller with respect to inclusion) with the left-side technique,
$\supsetneq$ the percentage of invariants stronger with the right-side technique,
and ``uncomparable'' gives the percentage of invariants that are uncomparable, i.e
neither greater nor smaller;
the code points where both invariants are equal make up the remaining percentage. See Tab.~\ref{tab:techniques} for details.}
  \label{fig:techniques}
\end {figure}

@@

We compared the precision of the different techniques described in this article as follows:
\begin{itemize}
\item For each program, we distinguished a set $P_R = P_W$ of suitable widening points by a simple algorithm: for each procedure, we compute the strongly connected components of its control-flow graph using Tarjan's algorithm; the targets of the back-edges of the depth-first search are added to~$P_R$. Note that the resulting set is not necessarily minimal, but is sufficient to disconnect all cycles.
(More sophisticated techniques are discussed in e.g. \citet{BourdonclePhd}.)

@@

Table \ref{tab:time} gives the execution time of the different analysis
techniques. It is interesting to see that \emph{Guided Static Analysis} and
\emph{Path-focusing} are sometimes faster than the classical
algorithm. The techniques we introduce in this article are slower, but speed is still on the same order of magnitude as the older ones.

@@

Statistically, the domain of convex polyhedra gives the better results, but
commonly yields greater invariants than the domains of octagons/intervals;
this is a known weakness of its widening operator.
\DM{Ici on pourrait citer Monniaux+Le Guen si c'etait publie!}
The Octagon domain appears to be much better than intervals; this is
unsurprising since in most programs and libraries, bounds on loop indices are
non constant: they depend on some parameters (array sizes etc.).

>>>>

\MM{Et que prend-t-on pour $\delta$ ? C'est 1 tout le temps ?}

@@

\MM{On peut peut-être citer le nom de l'outil (on garde PAGAI ?) ici.
  L'intérêt d'avoir un nom, c'est que les gens peuvent en parler plus
  facilement, et c'est plus facile de retrouver le papier origine d'un
  outil quand il cite son nom}

@@

SSA introduces $\phi$-functions at the head of a control code to define variables whose value depends on which incoming edge to this control node was last taken. For instance, for \lstinline|if (...) { x = 2*x+1; } else { x= 0; }|, then $x_2$ is defined as $\phi(2x_1+1,0)$. In this framework, each variable is uniquely defined as an arithmetic ($+$, $-$, $\times$, $/$) function of other variables that themselves may not be representable as affine linear functions, because they are defined using $\phi$-functions, bitwise arithmetic, loads from memory, or return values from function calls.

@@

\begin{figure}[h]
  \MM{Je trouve que les zéro-virgule-quelque-chose comme .17 font
    moches à côté des zéros, vu que les tableaux sont alignés sur le
    ., les ``0'' sont tout à gauche et les ``.17'' tout à droite. Je
    vote pour des 0.17 pour que l'alignement soit bon, ça rendrait la
    lecture plus facile je trouve}
  \begin{center}
    \input{./techniques.tex}
  \end{center} 
  \vspace{-20px}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).
  The $\subsetneq$ bars gives the percentage of invariants stronger (more precise; smaller with respect to inclusion) with the left-side technique,
$\supsetneq$ the percentage of invariants stronger with the right-side technique,
and ``uncomparable'' gives the percentage of invariants that are uncomparable, i.e
neither greater nor smaller;
the code points where both invariants are equal make up the remaining percentage. See Tab.~\ref{tab:techniques} for details.}
  \label{fig:techniques}
\end {figure}

@@

We compared the precision of the different techniques described in this article as follows:
\begin{itemize}
\item For each program, we distinguished a set $P_R = P_W$ of suitable widening points by a simple algorithm: for each procedure, we compute the strongly connected components of its control-flow graph using Tarjan's algorithm; the targets of the back-edges of the depth-first search are added to~$P_R$. Note that the resulting set is not necessarily minimal, but is sufficient to disconnect all cycles
(More sophisticated techniques are discussed in e.g. \citet{BourdonclePhd}).
\MM{Pourquoi cet item se trouve dans la section ``Precision of the
  various techniques'' et pas dans la section précédente sur
  l'implem ?}

@@

Table \ref{tab:time} gives the execution time of the different analysis
techniques. It is interesting to see that \emph{Guided Static Analysis} and
\emph{Path-focusing} are sometimes faster than the classical
algorithm.
\MM{Si y'a la place, on peut dire un truc du style: ``This can be
  explained by the fact that these algorithms start on a small subset
  of the graph, and then may need fewer iterations on the complete
  graph.''}
The techniques we introduce in this article are slower, but speed is still on the same order of magnitude as the older ones.

@@

Statistically, the domain of convex polyhedra gives the better results, but
commonly yields weaker invariants than the domains of octagons/intervals;
this is a known weakness of its widening operator.
\DM{Ici on pourrait citer Monniaux+Le Guen si c'etait publie!}
\MM{Oui, ça fait un peu [citation needed] comme phrase ;-)}
The Octagon domain appears to be much better than intervals; this is
unsurprising since in most programs and libraries, bounds on loop indices are
non constant: they depend on some parameters (array sizes etc.).
